---
title: "Pooling methods in Pairwise70"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pooling methods in Pairwise70}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
set.seed(42)
```

## Why another pooling estimator?

Standard random-effects meta-analysis assigns each study a weight
proportional to `1 / (vi + tau^2)`. That is optimal under the standard
assumptions but it doesn't ask the question that matters most for a clinical
reader: *if I removed one study, would my conclusion change?* Two studies
can have identical sampling variance but very different leverage on the
direction or significance of the pooled estimate.

`Pairwise70` provides six estimators that complement `metafor::rma()` by
asking variants of that question.

## A working dataset

```{r setup}
library(Pairwise70)

set.seed(42)
k <- 12
yi <- 0.30 + rnorm(k, sd = 0.15) + rnorm(k, sd = 0.20)
vi <- runif(k, 0.02, 0.20)
```

This is a small synthetic dataset on the log-effect-size scale (`k = 12`,
between-study SD `tau = 0.15`, varying sampling variances). It is large
enough that every method has something to do.

## A side-by-side comparison

`compare_pooling_methods()` runs the standard REML model, HKSJ, RVE (if
`clubSandwich` is installed) and each advanced method on the same data, so
you can read off how much each contributes:

```{r compare}
compare_pooling_methods(yi, vi)
```

If `RoBMA` is installed and you have time to wait, pass `include_robma =
TRUE` to add a Bayesian model-averaged row.

## MAFI-weighted meta-analysis (MWM)

MWM blends inverse-variance weights with leave-one-out *stability* weights.
A study whose removal would flip the direction or significance of the
pooled estimate gets a smaller stability weight, so it pulls the pooled
estimate less than a study that only adds precision.

```{r mwm}
mwm <- mafi_weighted_ma(yi, vi)
mwm$estimate
mwm$adjustment              # shift relative to REML
mwm$study_stability_scores  # per study; 1.0 = perfectly stable
```

The mixing constant is exposed as `stability_weight`. Setting it to `0`
recovers a precision-only estimator; `1` recovers a stability-only one.

## Adaptive robust pooling (ARP)

ARP fits the model with four estimators (REML, DL, PM, HKSJ) and combines
them with weights that depend on observed `I^2` and `k`. The standard error
includes a between-estimator term, so model uncertainty is reflected in the
CI.

```{r arp}
arp <- adaptive_robust_pooling(yi, vi)
arp$estimator_weights
arp$var_within     # average within-estimator variance
arp$var_between    # disagreement across estimators
```

The relative magnitudes of `var_within` and `var_between` tell you how much
of the uncertainty is from sampling vs. from not knowing which estimator
to trust.

## Sequential influence trimming (SIT)

SIT computes Cook's distance for each study, downweights any study above
threshold by `trim_factor`, refits, and repeats. The procedure converges
when influence drops below threshold or the pooled estimate stops moving.

```{r sit}
sit <- sequential_influence_trimming(yi, vi, influence_threshold = 0.5)
sit$iterations
sit$n_trimmed
sit$estimate_history       # trajectory across iterations
```

Unlike trim-and-fill, no studies are removed — weights only shrink — and
the final weights are part of the return value so the trimming is fully
auditable.

## Unified bias-stability framework (UBSF)

UBSF combines a *bias* adjustment (trim-and-fill plus an Egger flag) with a
*stability* adjustment (shrink toward the leave-one-out median) and inflates
the standard error when fragility is detected.

```{r ubsf}
ubsf <- unified_bias_stability(yi, vi)
ubsf$bias_detected
ubsf$direction_fragile
ubsf$se_inflation
ubsf$bias_adjustment
ubsf$stability_adjustment
```

`bias_weight` and `stability_weight` control how aggressive the two
adjustments are.

## Ensemble (EMA)

`ensemble_meta_analysis()` averages REML, MWM, ARP, SIT and UBSF (and
optionally RVE and RoBMA) with fixed weights, and reports both
within-method variance and between-method variance:

```{r ema}
ema <- ensemble_meta_analysis(yi, vi, include_rve = FALSE)
ema$method_weights
ema$agreement   # 1 - sd(estimates) / |ensemble estimate|
```

`agreement` is a useful one-glance check: high agreement means every method
points the same direction with similar magnitude; low agreement means the
choice of method materially changes the answer, which is itself a finding.

## When to reach for which

| Question | Try |
| --- | --- |
| "Is one study driving the conclusion?" | `mafi_weighted_ma()`, `sequential_influence_trimming()` |
| "Does my answer depend on which heterogeneity estimator I picked?" | `adaptive_robust_pooling()` |
| "Is there evidence of publication bias *and* fragility?" | `unified_bias_stability()` |
| "Reviewer asked for a sensitivity table" | `compare_pooling_methods()` |
| "I want one number that hedges across all of the above" | `ensemble_meta_analysis()` |

## Further reading

- The `analysis/` directory of this repository contains the full simulation
  and validation pipelines the methods were developed under.
- See `NEWS.md` for changes between releases and `CONTRIBUTING.md` for how
  to add a new method.
